---
title: "HW_9_SDS315"
output:
  pdf_document: default
  html_document: default
date: "2024-04-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
library(dplyr)
library(tidyverse)
library(mosaic)
library(ggplot2)
library(MatchIt)
library(effectsize)
library(kableExtra)
```

# HW 9

Name: Madeline Boss

EID: mrb5727

To see the code visit [Github](https://github.com/MadelineRBoss/HW9_SDS315)

## Probelm 1

```{r 1data, include=FALSE}
turnout_data <- read_csv("turnout.csv")
```
### Part A

```{r 1A, echo=FALSE}
prop_vote_GOTV_table = xtabs(~GOTV_call + voted1998, turnout_data)
prop_vote_GOTV_yes = prop_vote_GOTV_table[2,2]/(prop_vote_GOTV_table[2,2] + prop_vote_GOTV_table[2,1])

prop_vote_GOTV_no = prop_vote_GOTV_table[1,2]/(prop_vote_GOTV_table[1,2] + prop_vote_GOTV_table[1,1])

turnout_diffmean <- do(1000)*diffprop(voted1998 == 1 ~ GOTV_call, data=mosaic::resample(turnout_data))

turnout_diffmean_ci <- confint(turnout_diffmean, level = 0.95)
```

Proportion of those that received a GOTV call and voted is `r round(prop_vote_GOTV_yes, 2)`

The sample proportion of those who did not receive a GOTV call who voted in 1998 is `r round(prop_vote_GOTV_no, 2)`

The 95% confidence interval for the difference in proportions in voting for those who did and didn't get a call is `r round(turnout_diffmean_ci$lower, 2)` - `r round(turnout_diffmean_ci$upper, 2)`S

### Part B

#### Voting in 1996
```{r 2a_voted1996, echo=FALSE}
ggplot(turnout_data) + geom_bar(aes(x = factor(GOTV_call), y= voted1998, fill=factor(voted1996)), stat = "identity") + facet_wrap(~voted1996) + labs(x = "Got GOTV Call", y = "Count of People who Voted in 1998", title ="Voting in 1996 effect on getting GOTV call")
```

As seen in the graphs above, the amount of GOTV calls received for people who voted in 1996 was significantly higher. This makes it a confounding factor.



#### Major Party Membership
```{r 2a_majorpty, echo=FALSE}
ggplot(turnout_data) + geom_bar(aes(x = factor(GOTV_call), y= voted1998, fill=factor(MAJORPTY)), stat = "identity") + facet_wrap(~MAJORPTY) + labs(x = "Got GOTV Call", y= "Count of People who Voted in 1998" ,title ="Party membership effect on getting GOTV call")
```

As seen in the graph above, people who were affiliated with a political party were much more likely not to get a GOTV call. This makes party affliation a confunding factor.



#### Age
```{r 2A_AGE, echo= FALSE}
ggplot(turnout_data) + geom_boxplot(aes(group = factor(GOTV_call), AGE, fill = factor(voted1998))) + facet_wrap(~voted1998) + labs(title = "Average Age by if they Voted in 1998 and if they got a GOTV Call", y = "If they got a GOTV Call")
```

As seen in the graph above, people who are older were more likely to vote in 1998. Do to this, age is also a confounding factor.



### Part C
```{r 2C_math, echo = FALSE}
#Find macthed data
GOTV_match = matchit(GOTV_call ~ AGE + MAJORPTY + voted1996, data = turnout_data, ratio = 5)

GOTV_matches = match.data(GOTV_match)
```

#### Proof

```{r 2C_proof, echo=FALSE}
ggplot(GOTV_matches) + geom_bar(aes(x = factor(GOTV_call), y= voted1998, fill=factor(voted1996)), stat = "identity") + facet_wrap(~voted1996) + labs(x = "Got GOTV Call", y = "Count of People who Voted in 1998", title ="Voting in 1996 effect on getting GOTV call")

ggplot(GOTV_matches) + geom_bar(aes(x = factor(GOTV_call), y= voted1998, fill=factor(MAJORPTY)), stat = "identity") + facet_wrap(~MAJORPTY) + labs(x = "Got GOTV Call", y= "Count of People who Voted in 1998" ,title ="Party membership effect on getting GOTV call")

ggplot(GOTV_matches) + geom_boxplot(aes(group = factor(GOTV_call), AGE, fill = factor(voted1998))) + facet_wrap(~voted1998) + labs(title = "Average Age by if they Voted in 1998 and if they got a GOTV Call", y = "If they got a GOTV Call")
```

As seen above, after matching the GOTV call ratios are now approximately equal in 1996 voting and majority party, and the age boxplots among GOTV call is approximately equal.

```{r 2C_PartA, echo = FALSE}
#Find Part A stuff
prop_vote_GOTV_match_table = xtabs(~GOTV_call + voted1998, turnout_data)
prop_vote_GOTV_match_yes = prop_vote_GOTV_match_table[2,2]/(prop_vote_GOTV_match_table[2,2] + prop_vote_GOTV_match_table[2,1])

prop_vote_GOTV_match_no = prop_vote_GOTV_match_table[1,2]/(prop_vote_GOTV_match_table[1,2] + prop_vote_GOTV_match_table[1,1])

boot_matched_GOTV <- do(1000)*diffprop(voted1998==1 ~ GOTV_call, data=mosaic::resample(GOTV_matches))

ci_matched_GOTV <- confint(boot_matched_GOTV, level = 0.95)
```




#### Matching Statistics 

After the matching, here's the new desprictive statistics

The proportion of those who received a GOTV call who voted is `r round(prop_vote_GOTV_match_yes, 2)`

The proportion of those who received a GOTV call who voted is `r round(prop_vote_GOTV_match_no, 2)`

Finally the confidence interval is `r ci_matched_GOTV$lower` - `r ci_matched_GOTV$upper`

## Probelm 2 Manufacturing flaws in circuit boards
```{r 2_setup, include = FALSE}
solder_data <- read_csv("solder.csv")
```

### Part A
```{r 2a, echo=FALSE}
ggplot(solder_data) + geom_boxplot(aes(x=Opening, y=skips)) + labs(title = "Opening v Skips")

ggplot(solder_data) + geom_boxplot(aes(x=Solder, y=skips)) + labs(title = "Solder Type v Skips")
```

Starting with Opening Type, as the opening size decreases, there's an associated increase in skips. Next with Solder Type, it appears that when a solder is thinner, it has an increased chance of having skips.



### Part B
```{r 2b, echo=FALSE}
lm_solder <- lm(skips ~ Opening + Solder + Opening:Solder, solder_data)

solder_values <- coef(lm_solder) %>%
  round(3)

solder_95 <- confint(lm_solder) %>%
  round(3)

solder_tibble <- cbind(solder_values, solder_95)

kable_classic_2((kbl(solder_tibble)))
```

### Part C
The intercept (thick and large) has a value of 0.393. This means the baseline skips for thick and large is 0.393 skips. OpeningM has a value of 2.407, which means the effect of skips of 2.407 skips in isolation. OpeningS has a value of 5.127, which means the effect of skips of 5.127 skips in isolation. OpeningM:SolderThin has a value of -0.740, which means when the opening is medium and the solder thickness is medium it had -0.740 effect on skip amount. Finally, OpeningS:SolderThin has a value of 9.653, which means when the opening is medium and the solder thickness is medium it had 9.653 effect on skip amount.

### Part D
If I had to suggest an opening size and solder thickness I would suggest medium soldering with a thin thickness. This is because OpeningM:SolderThin was the only value to have a negative value, indicating it to have the least amount effect on skips.